# Reproducible Research: Peer Assessment 1
### Ashwin Kadle

```{r, echo=FALSE, results='hide', warning=FALSE}
library(ggplot2)
```

## Loading and preprocessing the data

#### 1. Code for reading in the dataset and/or processing the data

```{r}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
ActivityData <- read.csv('activity.csv')
```

#### What is mean total number of steps taken per day?

```{r}
StepsTakenPerDay <- tapply(ActivityData$steps, ActivityData$date, sum, na.rm=T)
```

#### 2. Histogram of the total number of steps taken each day

```{r}
qplot(StepsTakenPerDay, xlab='Total steps per day', ylab='Frequency')
```

#### 3. Mean and median number of steps taken each day

```{r}
StepsMean <- mean(StepsTakenPerDay)
StepsMedian <- median(StepsTakenPerDay)
```

* Mean steps per day    : `r StepsMean`
* Median steps per day  : `r StepsMedian`

## What is the average daily activity pattern?

```{r}
AverageStepsPerTimeBlock <- aggregate(x=list(meanSteps=ActivityData$steps), by=list(interval=ActivityData$interval), FUN=mean, na.rm=TRUE)
```

#### 1. Time series plot of the average number of steps taken

```{r}
ggplot(data=AverageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) + geom_line() + xlab("5-minute interval") + ylab("Average number of steps taken") 
```

#### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps? 

```{r}
MaxSteps <- which.max(AverageStepsPerTimeBlock$meanSteps)
TimeMaxSteps <-  gsub("([0-9]{1,2})([0-9]{2})","\\1:\\2", AverageStepsPerTimeBlock[MaxSteps,'interval'])
```

* Most Steps at: `r TimeMaxSteps`

## Imputing missing values

#### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
NAValues <- sum(is.na(ActivityData))

```

* Number of missing values: `r NAValues`

#### 2. Devise a strategy for filling in all of the missing values in the dataset.

My strategy is to use the mean for that 5-minute interval to fill each NA value in the steps column

#### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
NewActivityData <- ActivityData 

for (i in 1:nrow(NewActivityData)) {
  if (is.na(NewActivityData$steps[i])){
    NewActivityData$steps[i] <- AverageStepsPerTimeBlock[which(NewActivityData$interval[i]==AverageStepsPerTimeBlock$interval), ]$meanSteps
  }
}

NewData <- sum(is.na(NewActivityData))

```
* There are `r NewData` NA's in NewActivityData

#### 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r}
Su <- tapply(NewActivityData$steps, NewActivityData$date, sum, na.rm=T)
hist(Su, xlab = "sum of steps per day", main = "histogram of steps per day")
```

#### Do these values differ from the estimates from the first part of the assignment?

```{r}
newTotalSteps <- aggregate(NewActivityData$steps, 
                           list(Date = NewActivityData$date), 
                           FUN = "sum")$x
newMean <- mean(newTotalSteps)

newMedian <- median(newTotalSteps)



```
* New Mean of Steps   : `r newMean` 
* New Median of Steps : `r newMedian`

#### What is the impact of imputing missing data on the estimates of the total daily number of steps?

* Old mean `r StepsMean` : New mean  `r newMean`
* Old median `r StepsMedian` : New median  `r newMedian`

## Are there differences in activity patterns between weekdays and weekends?

#### 1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}


NewActivityData$weekday <- c("weekday")
NewActivityData[weekdays(as.Date(NewActivityData[, 2])) %in% c("Saturday", "Sunday", "saturday", "sunday"), ][4] <- 
c("weekend")
table(NewActivityData$weekday == "weekend")

NewActivityData$weekday <- factor(NewActivityData$weekday)
```

#### 2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

In order to visualize the difference bewteen weekends and days of the week, a new dataframe is created to be usable by the lattice package. First, 
the data are calculated:

```{r}
NewActivityData_weekend <- subset(NewActivityData, NewActivityData$weekday == "weekend")
NewActivityData_weekday <- subset(NewActivityData, NewActivityData$weekday == "weekday")

mean_NewActivityData_weekday <- tapply(NewActivityData_weekday$steps, NewActivityData_weekday$interval, mean)
mean_NewActivityData_weekend <- tapply(NewActivityData_weekend$steps, NewActivityData_weekend$interval, mean)
```

Then the dataframe is prepared and the plot is. plotted !

```{r, warning=FALSE}
library(lattice)
df_weekday <- NULL
df_weekend <- NULL
df_final <- NULL
df_weekday <- data.frame(interval = unique(NewActivityData_weekday$interval), avg = as.numeric(mean_NewActivityData_weekday), day = rep("weekday", length
(mean_NewActivityData_weekday)))
df_weekend <- data.frame(interval = unique(NewActivityData_weekend$interval), avg = as.numeric(mean_NewActivityData_weekend), day = rep("weekend", length
(mean_NewActivityData_weekend)))
df_final <- rbind(df_weekday, df_weekend)
```


```{r}
xyplot(avg ~ interval | day, data = df_final, layout = c(1, 2), 
       type = "l", ylab = "Number of steps")

```

